The importance of a complete statistical description of dependence between variables for ecology and related fields

### 2019-04-06 Paper.Rmd
Discussion edited for vegetation related drought

<!--Maybe a para on how likely one is to get copula structure via the Moran mechanism - considers
work on copula structure in environmental variables. Maybe mention possible future work: 
study copula structure of environmental variables (such as temperature and precip) that are known 
to be important in ecology, including how such copula structure may be changing and whether it 
has a geography of the kind studied in Walter et al.-->
One may expect our Moran-like mechanism for non-N copula structure and tail dependence to 
operate commonly, because Moran effects are common [@Liebhold2004], and 
non-N and tail-dependent copula structures have often been found in prior analyses of 
environmental variables. If large meteorological events are also spatially widespread, then
one would expect environmental variables associated with these events to take extreme
values simultaneously across large spatial areas, producing tail dependence in measurements
made at different locations. Non-extreme values, not associated with extreme events, may
be more associated with local weather phenomena, and therefore 
less prone to being correlated across large areas.
Using a bivariate mixed model approach with copulas, 
@Serinaldi2008 examined the spatial dependence 
of rainfall observations (7 years with 30 minute time resolution) 
within a rain gauge network in Central Italy. Gumbel or Student 2-copulas 
were possible candidates for modelling bivariate dependence at several time scales, and 
neither of these is an N copula. @liu2016probabilistic used a copula 
based approach in assessing the probability of vegetation related drought across China. 
<!--***DAN: Shya, what do you mean by "vegetation related drought"? Is this a kind of drought
that is somehow caused by the vegetation? That's what it sounds like, and it sounds like what
they studied was the probability that such a drought occurred. If that's not what you mean, please
rephrase.
***Shya: I used the exact term "vegetation-related drought" they used in the paper @liu2016probabilistic, I think they 
wanted to mean that you can monitor the drought risk by tracking change in NDVI spatial greenness with precipitation or temperature (climatic variable), which can be considered as a proxy indicator of vegetation response to climate change. They also showed vegatation-drought stress analysis, which I also noticed in some other papers too. There are several studies using drought and NDVI data (mainly using correlation, regression) but they reported it as the first study using joint probability for drought-risk assessment. I have highlighted the info which may be important to us and sent you in mail.-->
In a majority of their cases, they found 
non-N copula structure in the dependence between a vegatation index 
and climatic variables (e.g., precipitation and temperature). 
Another long-term study (1950–2014) in the Loess Plateau of China [@She2018]
showed that a Gumbel copula (which has upper tail dependence) could effectively model 
the spatial dependence of drought variables 
(duration and severity) in the study area. Bivariate copula analysis was also used in forecasting the 
co-occurrence of extreme events (flood or drought) 
over the North Sikkim Himalayas, India, using spatial data sets at different 
time scales [@Goswami2018].

>>>>

<!--Maybe a para on how likely one is to get copula structure via the Moran mechanism - considers
work on copula structure in environmental variables. Maybe mention possible future work: 
study copula structure of environmental variables (such as temperature and precip) that are known 
to be important in ecology, including how such copula structure may be changing and whether it 
has a geography of the kind studied in Walter et al.-->
One may expect our Moran-like mechanism for non-N copula structure and tail dependence to 
operate commonly, because Moran effects are common [@Liebhold2004], and 
non-N and tail-dependent copula structures have often been found in prior analyses of 
environmental variables. If large meteorological events are also spatially widespread, then
one would expect environmental variables associated with these events to take extreme
values simultaneously across large spatial areas, producing tail dependence in measurements
made at different locations. Non-extreme values, not associated with extreme events, may
be more associated with local weather phenomena, and therefore 
less prone to being correlated across large areas.
Using a bivariate mixed model approach with copulas, 
@Serinaldi2008 examined the spatial dependence 
of rainfall observations (7 years with 30 minute time resolution) 
within a rain gauge network in Central Italy. Gumbel or Student 2-copulas 
were possible candidates for modelling bivariate dependence at several time scales, and 
neither of these is an N copula. @liu2016probabilistic used a copula 
based approach in assessing the drought risk with joint probability distribution 
between vegetation and climatic variables across China. 
In a majority of their cases, they found 
non-N copula structure in the dependence between a vegatation index 
and climatic variables (e.g., precipitation and temperature). 
Another long-term study (1950–2014) in the Loess Plateau of China [@She2018]
showed that a Gumbel copula (which has upper tail dependence) could effectively model 
the spatial dependence of drought variables 
(duration and severity) in the study area. Bivariate copula analysis was also used in forecasting the 
co-occurrence of extreme events (flood or drought) 
over the North Sikkim Himalayas, India, using spatial data sets at different 
time scales [@Goswami2018].
